Relative parabolic factorizations of Chevalley groups

### 2015-04-07 congrsubgr-boundgen.tex
BMS reference

\textsc{Case 2:} $\alpha<\beta$. Since $\mathbb{Z}[\sfrac{1}{p}]/I$ is finite, there exists $k>0$ such that $p^k\equiv 1\pmod I$. One can choose $k>\beta-\alpha$. Then $k+\alpha>-k+\beta$ and
\[ g\cdot h\left(p^k\right) =
\begin{pmatrix} p^\alpha a & p^\beta bm \\ * & * \end{pmatrix}
\begin{pmatrix} p^k & 0 \\ 0 & p^{-k} \end{pmatrix}=
\begin{pmatrix} p^{k+\alpha} a & p^{-k+\beta} bm \\ * & * \end{pmatrix}.
\]
One can apply Case 1 for the latter, so
$g=x_{21}z_{12}x_{12}x_{21}h^{-1}$,
and $x_{12}h^{-1}$ can be processed in the same way, as in Lemma \ref{lemma:srRI1}:
$g=x_{21}z_{12}x_{12}\cdot z_{12}x_{21}x_{12}$.
\end{proof}
Let $\mathcal{O}_S$ be a Dedekind ring of arithmetic type, defined by a finite set $S$ of places of some global field, and $I$ an ideal in $\mathcal{O}_S$. It is shown in \cite{TavgenThesis} that $E(\Phi,I)$ has finite width with respect to $\{x_\alpha(s)\mid\alpha\in\Phi,\ s\in I\}$. From this one immediately gets the bounded generation for $\G(\Phi,R,I)$.

Assume for simplicity that $\K_1(\Phi,\mathcal{O}_S,I)=1$.
\begin{lemma}
If $\rk\Phi\geqslant2$, then $\G(\Phi,\mathcal{O}_S,I)$ has finite width with respect to $z_\alpha$.
\end{lemma}
\begin{proof}
Because of the Bass---Kolster decomposition it suffices to express every element $g$ of $\SL(2,\mathcal{O}_S,I)$ as a product of a bounded number of generators in a group of rank $2$. Write
\[ g=\begin{pmatrix}1+a & b \\ c & 1+d\end{pmatrix}\in\SL(2,\mathcal{O}_S,I),\quad a,b,c,d\in I. \]

>>>>

\textsc{Case 2:} $\alpha<\beta$. Since $\mathbb{Z}[\sfrac{1}{p}]/I$ is finite, there exists $k>0$ such that $p^k\equiv 1\pmod I$. One can choose $k>\beta-\alpha$. Then $k+\alpha>-k+\beta$ and
\[ g\cdot h\left(p^k\right) =
\begin{pmatrix} p^\alpha a & p^\beta bm \\ * & * \end{pmatrix}
\begin{pmatrix} p^k & 0 \\ 0 & p^{-k} \end{pmatrix}=
\begin{pmatrix} p^{k+\alpha} a & p^{-k+\beta} bm \\ * & * \end{pmatrix}.
\]
One can apply Case 1 for the latter, so
$g=x_{21}z_{12}x_{12}x_{21}h^{-1}$,
and $x_{12}h^{-1}$ can be processed in the same way, as in Lemma \ref{lemma:srRI1}:
$g=x_{21}z_{12}x_{12}\cdot z_{12}x_{21}x_{12}$.
\end{proof}
Let $\mathcal{O}_S$ be a Dedekind ring of arithmetic type, defined by a finite set $S$ of places of some global field $k$, and $I$ an ideal in $\mathcal{O}_S$. In case $k$ has a real embedding, then $\K_1(\Phi,\mathcal{O}_S,I)=1$ for any ideal $I\trianglelefteq\mathcal{O}_S$ and any root system $\Phi$ of rank $\geqslant2$ (see~\cite{BassMilnorSerre}). It is shown in \cite{TavgenThesis} that $E(\Phi,I)$ has finite width with respect to $\{x_\alpha(s)\mid\alpha\in\Phi,\ s\in I\}$. From this one immediately gets the bounded generation for $\G(\Phi,R,I)$.
\begin{lemma}
If $\rk\Phi\geqslant2$ and $k$ has a real embedding, then $\G(\Phi,\mathcal{O}_S,I)$ has finite width with respect to $z_\alpha$.
\end{lemma}
\begin{proof}
Because of the Bass---Kolster decomposition it suffices to express every element $g$ of $\SL(2,\mathcal{O}_S,I)$ as a product of a bounded number of generators in a group of rank $2$. Write
\[ g=\begin{pmatrix}1+a & b \\ c & 1+d\end{pmatrix}\in\SL(2,\mathcal{O}_S,I),\quad a,b,c,d\in I. \]
